\textls[-0]{\thetitle}

### 2023-03-07 metatheory-condensed.tex
Working on examples.
Combine Wtype into the Figure.

\begin{figure}
  \begin{tabular}{
    >{\columncolor[HTML]{C0C0C0}}l lcl}
    \hline
    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}surface syntax} &
      \multicolumn{1}{l|}{Equivalent Type} &
      \multicolumn{1}{l}{} &
      \multicolumn{1}{l|}{} \\
    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}\lsti!| unit : tm!} &
      \multicolumn{1}{l|}{$\bbone \times (\bbzero \to tm) \to tm$} &
      $\tau'$ &
      \multicolumn{1}{l|}{$\coloneqq\WSigAdd{\WSigEmp}{\bbone}{\bbzero}$} \\
    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}\lsti!| app  : tm -> tm -> tm!} &
      \multicolumn{1}{l|}{$\bbone \times (\bbtwo \to tm) \to tm$} &
      $\tau$ &
      \multicolumn{1}{l|}{\cellcolor[HTML]{FFFFC7}$\coloneqq\WSigAdd{\tau'}{\bbone}{\bbtwo}$} \\ \hline
    \multicolumn{1}{c}{\cellcolor[HTML]{C0C0C0}\lsti!unit!} &
      $\coloneqq \Wsup{1}{\tau}{\unit}{"elim-"\bbzero}$ &
      \cellcolor[HTML]{C0C0C0}\lsti!app unit unit! &
      $\coloneqq\Wsup{0}{\tau}{\unit}{"unit"}$
    \end{tabular}
  \caption{W-type encoding of "tm \{unit, app\}"}
  \label{fig:Wtype-example}
\end{figure}

We illustrate one concrete example in \cref{fig:Wtype-example} of a "tm" with only two constructors, 
also as a review for how W-type encodes a given inductive type. The key insight 
is, to encode one constructor into a pair $(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})$, the non-recursive part will be stored in $\wsigproj{i}{1}{\wsig}$ 
while the ``arity'' of the  recursive part stores in $\wsigproj{i}{2}{\wsig}$. 
For example, $"app" : ("tm" \to "tm" \to "tm")$ 
$\cong ((\bbtwo \to "tm") \to "tm")$, saying the ``arity'' of the recursive part is $\bbtwo$ 
(or boolean type, as a type with two elements). Similarly, the ``arity'' of recursive part of "unit" is 0. 
Both constructors don't store other information thus $\bbone$. We hightlight $\tau$ as the 
signature of our "tm" and showcase the meta-theoretic encoding of "unit" and "app~unit~unit". $"elim-"\bbzero$ indicates the elimination of the bottom type $\bbzero$ in the context. Generally speaking, the expressiveness of W-types comes from the ``arity'', as it encodes the case when recursive structure stores as the output of a function, mirroring the strict positivity of Inductive type.

@@

\item 
The next four fields model the four constructors of~\lsti{tm}.
Constructor \lsti{tm_unit} is modeled as $\Wsup{}{\wsig_{\texttt{tm}}}{\top}{\bot}$
and typed with $\El{\mathit{self}.\texttt{tm}}$, where $\mathit{self}$ stands for
the typing context containing the first field~\lsti{tm}.
The W-type signature $\wsig_{\texttt{tm}}$ is exposed in
this typing context; that is, \lsti{tm} has type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$.
So $\El{\mathit{self}.\texttt{tm}}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
as required by rule \ruleref{tm/wsup}.

>>>>

% \begin{figure}
%   \begin{tabular}{llll}
%     \hline
%     \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}surface syntax} &
%       \multicolumn{1}{l|}{Equivalent Type} &
%        &
%       \multicolumn{1}{l|}{} \\
%     \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}| unit : tm} &
%       \multicolumn{1}{l|}{$\bbone \times (\bbzero \to tm) \to tm$} &
%       \multicolumn{1}{c}{$\tau'$} &
%       \multicolumn{1}{l|}{$\coloneqq\WSigAdd{\WSigEmp}{\bbone}{\bbzero}$} \\
%     \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}| app  : tm -\textgreater tm -\textgreater tm} &
%       \multicolumn{1}{l|}{$\bbone \times (\bbtwo \to tm) \to tm$} &
%       \multicolumn{1}{c}{$\tau$} &
%       \multicolumn{1}{l|}{\cellcolor[HTML]{FFFFC7}$\coloneqq\WSigAdd{\tau'}{\bbone}{\bbtwo}$} \\ \hline
%     \multicolumn{1}{c}{\cellcolor[HTML]{C0C0C0}unit} &
%       $\coloneqq \Wsup{1}{\tau}{\unit}{"elim-"\bbzero}$ &
%       \multicolumn{1}{c}{\cellcolor[HTML]{C0C0C0}app unit unit} &
%       $\coloneqq\Wsup{0}{\tau}{\unit}{"unit"}$ \\ \hline
%     \cellcolor[HTML]{ECF4FF}$\lsig' = \LSigAdd{\LSigEmp}{\pi_2}{\TyS{\wcode{\tau'}}}$ &
%       $o' = \LkgAdd{\LkgEmp}{\wcode{\tau'}}$ &
%       \cellcolor[HTML]{ECF4FF}$\lsig = \LSigAdd{\LSigEmp}{\pi_2}{\TyS{\wcode{\tau}}}$ &
%       $o = \LkgAdd{\LkgEmp}{\wcode{\tau}}$ \\
%     \cellcolor[HTML]{ECF4FF}$\lsig'^+ = \LSigAdd{\lsig'}{s'}{..}$ &
%       $o'^+ = \LkgAdd{o'}{t}$ &
%       \cellcolor[HTML]{ECF4FF}$\lsig^+ = \LSigAdd{\lsig}{s}{..}$ &
%       $o^+ = \LkgAdd{o}{\lkgproj{2}{o'^+}}$ \\
%     \multicolumn{2}{l}{$A = \TySigma{\cU}{\El{\pi_2} \to \El{\pi_2}} $} &
%       \multicolumn{2}{l}{$t = \goodTerm{\cdot , A}{\snd{\pi_2}}{\El{\fst{\pi_2}} \to \El{\fst{\pi_2}} }$} \\
%     $\goodTerm{\cdot , \TyTkg{\lsig'}}{s'}{\sub{A}{\pi_1}}$ &
%        &
%       $\goodTerm{\cdot , \TyTkg{\lsig}}{s}{\sub{A}{\pi_1}}$ &

%     \end{tabular}
%   \caption{W-type encoding of "tm \{unit, app\}"}
%   \label{fig:Wtype-example}
% \end{figure}

% We illustrate one concrete example in \cref{fig:Wtype-example} of a "tm" with only two constructors, 
% also as a review for how W-type encodes a given inductive type. The key insight 
% is, to encode one constructor into a pair $(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})$, the non-recursive part will be stored in $\wsigproj{i}{1}{\wsig}$ 
% while the ``arity'' of the  recursive part stores in $\wsigproj{i}{2}{\wsig}$. 
% For example, $"app" : ("tm" \to "tm" \to "tm")$ 
% $\cong ((\bbtwo \to "tm") \to "tm")$, saying the ``arity'' of the recursive part is $\bbtwo$ 
% (or boolean type, as a type with two elements). Similarly, the ``arity'' of recursive part of "unit" is 0. 
% Both constructors don't store other information thus $\bbone$. We hightlight $\tau$ as the 
% signature of our "tm" and showcase the meta-theoretic encoding of "unit" and "app~unit~unit". $"elim-"\bbzero$ indicates the elimination of the bottom type $\bbzero$ in the context. Generally speaking, the expressiveness of W-types comes from the ``arity'', as it encodes the case when recursive structure stores as the output of a function, mirroring the strict positivity of Inductive type.

@@

\item We illustrate the structure of W-type signature (as a list) here. $\wsig_{\texttt{tm}}^0, \wsig_{\texttt{tm}}^1$ are (the subparts of) the signatures of constructors \{"tm\_unit"\} and \{"tm\_unit","tm\_app"\} respectively. Each item in this W-type signature, having two types, corresponds to one constructor, where the first type stores the non-recursive part of the constructor and the second type stores the ``arity'' of the  recursive part. For example, the type of $"app" : ("tm" \to "tm" \to "tm")$ $\cong (\bbone \times (\bbtwo \to "tm") \to "tm")$, corresponds to the premise of rule \ruleref{tm/wsup} (where $\wsigproj{0}{1}{\wsig_{\texttt{tm}}^1} = \bbone, \wsigproj{0}{2}{\wsig_{\texttt{tm}}^1} = \bbtwo$). This type means there is no information on non-recursive part (thus use unit type $\bbone$) and the ``arity'' of the recursive part is $\bbtwo$ and thus $\wsig_{\texttt{tm}}^1 = \WSigAdd{\wsig_{\texttt{tm}}^0}{\bbone}{\bbtwo}$.  The story is similar for $\wsig_{\texttt{tm}}^0$, where there is no recursive part at all and thus the ``arity'' is zero.

Generally speaking, the expressiveness of W-types comes from the ``arity'', as it encodes the case when recursive structure stores as the output of a function, mirroring the strict positivity of Inductive type.
\item 
The next four fields model the four constructors of~\lsti{tm}.
Constructor \lsti{tm_unit} is modeled as $\Wsup{}{\wsig_{\texttt{tm}}}{\top}{\bot}$
and typed with $\El{\mathit{self}.\texttt{tm}}$, where $\mathit{self}$ stands for
the typing context containing the first field~\lsti{tm}.
The W-type signature $\wsig_{\texttt{tm}}$ is exposed in
this typing context; that is, \lsti{tm} has type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$.
So $\El{\mathit{self}.\texttt{tm}}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
as required by rule \ruleref{tm/wsup}.
